---
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=1.5cm"
output:
  pdf_document:
    toc: false
header-includes:
 - \usepackage{helvet}
 - \renewcommand{\familydefault}{\sfdefault}
---

<style>
body { font-family: Helvetica, sans-serif; margin: 20px; }
.blue-panel { background-color: #e8f4f8; border: 2px solid #4a90e2; padding: 20px; margin: 20px 0; border-radius: 5px; }
h1, h2, h3 { color: #2c3e50; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "100%",
  dev = "png",        # force PNG instead of SVG (SVG is huge for rasters)
  dpi = 300
)

library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(scales)
library(gridExtra)
library(grid)
library(ggtext)
library(lubridate)
library(sf)
library(patchwork)
# Additional libraries for weather data
library(httr)
library(jsonlite)

# Set data path
data_path <- '/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/'

## Get station metadata
meta <- readxl::read_xlsx(paste0(data_path,"MetaData.xlsx"))

## Set station of interest
station <- 21

# Load spatial data
# Land polygon (coastline)
land_sf <- read_sf('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/GIS/Land/gadm36_0.shp')

#PDAs polygon
pda_sf <- read_sf('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/GIS/PDAs/PDAs.shp')

# Stations point data (8 stations in a grid)
stations_sf <- read_sf("/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/GIS/s3_locations_new_wgs84.shp")
names(stations_sf)[1]<-"Station"

# Get coordinates of the focal station
focal_station <- stations_sf %>% filter(Station == station)

# Define bounding box (if not provided, use extent of all data)
bbox <- c(xmin = -8, xmax = -4, ymin = 49.8, ymax = 52.2)

```

```{r graphics, include=FALSE}

theme_custom <- theme_classic()+
  theme(
    axis.line = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.text = element_text(family = "Helvetica", size = 10),
    axis.title = element_text(family = "Helvetica", size = 12),
    panel.background = element_rect(fill = "white", colour = "NA"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.grid = element_line(colour = "grey90", linewidth = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.direction = "horizontal",
    plot.title = element_markdown(family = "Helvetica", color = "black", size = 16)
  )

```

# Station S`r sprintf("%02d",station)` - Service Report

```{r title_page_layout, fig.width=12, fig.height=6, out.width="100%"}

# Create a simple title page with station info on left and placeholder for map on right
start <- meta$DateTime_deploy_UTC[meta$Station==paste0("S",station)]
end <- meta$DateTime_retrieve_UTC[meta$Station==paste0("S",station)]

# Station details
station_info <- data.frame(
  label = c("Station N:",
            "Deployment period:",
            "Location (WGS84):",
            "Depth:",
            "VR2AR serial:",
            "FPOD serial:",
            "Hydrophone serial:",
            "Hydrophone duty cycle:",
            "Report generated"),
  value = c(paste0("S", sprintf("%02d",station)),
            paste0(date(start)," to ", date(end)),
            paste0(meta$Latitude_deploy[meta$Station==paste0("S", sprintf("%02d",station))],"°,  ",meta$Longitude_deploy[meta$Station==paste0("S", sprintf("%02d",station))],"°"),
            "120 m",
            meta$ARC_VR2AR[meta$Station==paste0("S", sprintf("%02d",station))],
            meta$FPOD[meta$Station==paste0("S", sprintf("%02d",station))],
            meta$Hydrophone[meta$Station==paste0("S", sprintf("%02d",station))],
            "50% on (15 min in 30 min)",
            format(Sys.Date(), "%B %d, %Y")
  )
)

# Create text layout (left side)
left_plot <- ggplot() +
  # geom_rect(aes(xmin = 0, xmax = 1, ymin = 0.82, ymax = 1.0),
  #           fill = "#85bec7", color = "black", linewidth = 0.5) +
  geom_rect(aes(xmin = 0, xmax = 2, ymin = 0, ymax = 1),
            fill = "white", color = "black", linewidth = 0.5) +
  # geom_text(aes(x = 0.1, y = 0.9, label = "Station Details"),
  #           size = 6, fontface = "bold", hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.93,
                label = paste(station_info$label[1], station_info$value[1], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.85,
                label = paste(station_info$label[9], station_info$value[9], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.72,
                label = paste(station_info$label[2], station_info$value[2], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.62,
                label = paste(station_info$label[3], station_info$value[3], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.52,
                label = paste(station_info$label[4], station_info$value[4], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.38,
                label = paste(station_info$label[5], station_info$value[5], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.28,
                label = paste(station_info$label[6], station_info$value[6], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.18,
                label = paste(station_info$label[7], station_info$value[7], sep = " ")),
            size = 5, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.08,
                label = paste(station_info$label[8], station_info$value[8], sep = " ")),
            size = 5, hjust = 0) +
  xlim(0, 2) + ylim(0, 1) +
  scale_x_continuous(expand = c(0,0), breaks = c(0,1), labels = c(" "," "))+
  scale_y_continuous(expand = c(0,0))+
  theme_void() +
  theme(
    axis.title.x = element_text(family = "Helvetica", size = 12),
    axis.text.x = element_text(family = "Helvetica", size = 10),
    panel.background = element_rect(fill = "white", colour = "NA"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    plot.title = element_text(family = "Helvetica", size = 16, face=("bold")),
    panel.margin = unit(0, 'mm')
  )+
  labs(title = "Station details", x = " ")

# Create placeholder map (right side) - simple scatter plot as placeholder
map_plot <- ggplot() +
  # Land
  geom_sf(data = land_sf, fill = "gray80", color = "gray50", size = 0.3) +
  #PDAs
  geom_sf(data = pda_sf, fill = "darkblue", colour = "darkblue", alpha=0.3)+
  # All stations (small points)
  geom_sf(data = stations_sf, shape = 21, fill = "white", color = "black", size = 2) +
  # Focal station (larger, different symbol)
  geom_sf(data = focal_station, shape = 21, fill = "red", color = "black", size = 4) +
  # Station label
  geom_sf_text(data = focal_station, aes(label = Station),
               nudge_x = 0.07, nudge_y = 0.07, size = 5, fontface = "bold") +
  # Crop to bounding box
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]),
           ylim = c(bbox["ymin"], bbox["ymax"]),
           expand = FALSE) +
  theme(
    axis.line = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.text = element_text(family = "Helvetica", size = 10),
    axis.title = element_text(family = "Helvetica", size = 12),
    
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "#e8f4f8", color = "black", linewidth = 1),
    panel.grid = element_line(colour = "grey90"),
    
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.direction = "horizontal",
    
    plot.title = element_text(family = "Helvetica", size = 16, face="bold"),
    plot.margin = margin(0, 0, 0, 0)
  )+
  labs(title = "Station location", x = "Longitude", y = "Latitude")

# Combine plots side by side
grid.arrange(left_plot, map_plot, ncol = 2, widths = c(0.4, 0.6))
```


## Environmental Conditions

```{r ctd_current_wind, fig.width=12, fig.height=8, out.width="100%"}

# Download weather data
base_url <- "https://archive-api.open-meteo.com/v1/archive"

# Build URL
url <- paste0(base_url,
              "?latitude=", focal_station$Y,
              "&longitude=", focal_station$X,
              "&start_date=", as.character(date(start)),
              "&end_date=", as.character(date(end)),
              "&hourly=temperature_2m,precipitation,wind_speed_10m,wind_direction_10m",
              "&timezone=GMT")

# Make request
response <- GET(url)

if (status_code(response) == 200) {
  data <- fromJSON(content(response, "text"))

  # Create data frame
  weather_df <- data.frame(
    DateTime = as.POSIXct(data$hourly$time, format = "%Y-%m-%dT%H:%M",tz = "GMT"),
    temperature = data$hourly$temperature_2m,
    rainfall = data$hourly$precipitation,
    wind_speed = data$hourly$wind_speed_10m,
    wind_direction = data$hourly$wind_direction_10m
  )
}
# Summarise at 12 hour intervals
weather_df <- weather_df %>%
  mutate(DateTime = floor_date(DateTime, "12 hours")) %>%
  group_by(DateTime) %>%
  summarise(
    temperature_mean    = mean(temperature),
    temperature_sd      = sd(temperature),
    wind_speed_mean     = mean(wind_speed),
    wind_speed_sd       = sd(wind_speed),
    wind_direction_mean = mean(wind_direction),
    wind_direction_sd   = sd(wind_direction)
  )


# Plot Wind Speed data
wind_speed <- ggplot(data = weather_df, aes( x = DateTime, y = wind_speed_mean))+
  geom_ribbon(aes(ymin = 0, ymax = wind_speed_mean), fill = "#001744", colour = "NA", alpha = 0.7)+
  geom_line(colour = "#001744", linewidth = 0.5)+
  scale_y_continuous("Wind speed (m/s)", expand = c(0,0))+
  scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "<b>Wind speed</b> <i style='font-size:14pt'>(source: OpenMeteo)</i>")

# Plot wind direction data
wind_dir <- ggplot(data = weather_df, aes( x = DateTime, y = wind_direction_mean))+
  geom_line(colour = "#001744", linewidth = 0.5)+
  scale_y_continuous("Wind direction", expand = c(0,0))+
  scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "<b>Wind direction</b> <i style='font-size:14pt'>(source: OpenMeteo)</i>")

# Get current data from POLPRED
current_data <- read.csv('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/POLPRED/S21_14oct25_16Jan26.csv', skip = 4)
current_data$time <- as.POSIXct(current_data$Date.Time..UTC., format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")

current_data <- filter(current_data,
                       time>= start, time <= end)
names(current_data)[3] <- "current_speed"

cmean <- current_data %>%
  mutate(DateTime = floor_date(time, "12 hours")) %>%
  group_by(DateTime) %>%
  summarise(
    current_speed_mean    = mean(current_speed),
    current_speed_sd    = sd(current_speed)
  )

# Plot current data
curr_speed <- ggplot(current_data, aes(x = time, y = current_speed)) +
  geom_point(size = 1, colour = "#1b9e77", alpha = 0.5) +
  geom_ribbon(data= cmean, aes(x = DateTime, ymin = current_speed_mean-current_speed_sd, ymax = current_speed_mean+current_speed_sd), fill = "grey80", alpha = 0.5, inherit.aes = F)+
  geom_line(data= cmean, aes(x = DateTime, y = current_speed_mean), colour = "#1b9e77", linewidth = 1.5, inherit.aes = F)+
   scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "<b>Bottom current speed</b> <i style='font-size:14pt'>(source: Polpred C320)</i>")+
  labs(x = "Time (hours)", y = "Speed (m/s)")

# Get temeprature data from VR2AR
# Read skipping the first 4 header rows, no column names
raw <- read_csv("/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/Processed_Sensors/VR2AR/S21/S21_R1/VR2AR-69 556544 2026-02-13 155019 SENSORS.csv",
                skip = 4,
                col_names = FALSE,
                show_col_types = FALSE)

# Keep only TEMP rows and select relevant columns
temp_df <- raw %>%
  filter(X1 == "TEMP") %>%
  select(
    record_type  = X1,
    datetime     = X2,
    model        = X6,
    serial       = X7,
    temp_ambient = X8
  ) %>%
  mutate(
    datetime     = dmy_hm(datetime),
    temp_ambient = as.numeric(temp_ambient),
    DateTime = floor_date(datetime, "12 hours")) %>%
  filter(DateTime >= start,
         DateTime <= end)%>%
  group_by(DateTime) %>%
  summarise(
    temperature_mean    = mean(temp_ambient),
    temperature_sd      = sd(temp_ambient)
  )

t_pod <- read.csv('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/Processed_Sensors/FPod/S21/S21_R1/S21_R1 2025 10 10 FPOD_8432 file0 PART 93d 11h 38m mixed  DetEnv.csv')
t_pod$DateTime <- as.POSIXct(t_pod$ChunkEnd, formqat = "%Y/%m/%d %H:%M", tz = "GMT")
t_pod <- filter (t_pod, DateTime >= start, DateTime <= end)

temp <- ggplot(temp_df, aes(x = DateTime)) +
  geom_line(data = t_pod, aes(y = Temp), colour = "black", alpha = 0.8, linewidth = 1)+
  geom_line(aes(y = temperature_mean), colour = "#d17432", linewidth = 1)+
  #geom_ribbon(data = weather_df, aes(ymin = temperature_mean-temperature_sd, ymax = temperature_mean+temperature_sd), fill = "grey80", alpha = 0.5)+
  #geom_line(data = weather_df, aes(y = temperature_mean))+
   scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "<b>Bottom water temperature</b> <i style='font-size:14pt'>(source: VR2AR & FPod)</i>")+
  labs(x = "Date", y = "°C")+
  annotation_custom(
    grob = grobTree(
      segmentsGrob(x0 = 0.87, x1 = 0.89, y0 = 0.875, y1 = 0.875,
                   gp = gpar(col = "#d17432", lwd = 2)),
      textGrob("VR2AR", x = 0.895, y = 0.875, hjust = 0,
               gp = gpar(fontsize = 8, fontfamily = "Helvetica")),
      segmentsGrob(x0 = 0.94, x1 = 0.96, y0 = 0.875, y1 = 0.875,
                   gp = gpar(col = "black", lwd = 2)),
      textGrob("FPod", x = 0.965, y = 0.875, hjust = 0,
               gp = gpar(fontsize = 8, fontfamily = "Helvetica"))
    ),
    xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf
  )

grid.arrange(wind_speed, wind_dir, curr_speed, temp, ncol = 1)
```

\pagebreak

## Soundscape and Noise

\vspace{0.1in}

```{r quantiles_plot, fig.width=12, fig.height=4.5, out.width="100%"}
#quant_file <- file.path(output_path, "all_quantiles.csv")
quant_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/quantile_spectral_density_12h.csv"))

if (file.exists(quant_file)) {
  q <- read_csv(quant_file, show_col_types = FALSE)

  ggplot(q, aes(x = TOB, y = SPL, color = as.factor(percentile))) +
    geom_line(linewidth = 0.9) +
    scale_x_log10(
      breaks = c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000),
      labels = c("10","20","50","100","200","500","1k","2k","5k","10k","20k","50k"),
      expand = c(0,0)
    ) +
    scale_color_manual(
      values = c("1" = "#fee5d9", "5" = "#fcae91", "25" = "#fb6a4a",
                 "75" = "#de2d26", "95" = "#a50f15","99" = "darkred", "50" = "#006837"),
      labels = c("1" = "1st %ile", "5" = "5th %ile", "25" = "25th %ile", "50" = "Median",
                 "75" = "75th %ile", "95" = "95th %ile", "99" = "99th %ile")
    ) +
    theme_custom +
    theme(legend.position = "right", legend.direction = "vertical")+
    labs(title = "<b>Spectral Quantiles</b> <i style='font-size:14pt'>(Distribution of acoustic energy across TOB frequencies)</i>")+
    labs(x = "Frequency (Hz)", y = "Power (dB re 1µPa²/Hz)")
} else {
  cat("Quantiles file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/Quatiles.jpg", last_plot(), height = 10, width = 15, units = "cm")
```


\vspace{0.2in}

```{r ltsa_plot, fig.width=12, fig.height=7, out.width="100%"}
#ltsa_file <- file.path(output_path, "all_ltsa.csv")
ltsa_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/ltsa_12h.csv"))

if (file.exists(ltsa_file)) {
  ltsa <- read_csv(ltsa_file, guess_max = 10000, show_col_types = FALSE)

  ltsa$DateTime <- as.POSIXct(ltsa$DateTime, tz = "GMT")

  ggplot(ltsa, aes(x = DateTime, y = TOB, fill = SPL)) +
    geom_raster(interpolate =T) +
    geom_vline(xintercept=as.POSIXct(c("2025-11-01 00:00:00",
                                       "2025-12-01 00:00:00",
                                       "2026-01-01 00:00:00"), tz = "UTC"),
               colour = "white", linetype = 2)+
    scale_y_log10(limits = c(10,9400), expand = c(0,0),
                  breaks = c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 9000),
                  labels = c("10","20","50","100","200","500","1k","2k","5k","9k")) +
    scale_x_datetime(expand = c(0,0), breaks = "15 days", date_labels = "%b %d")+
    #scale_y_log10()+
    scale_fill_viridis_c(name = "Power\n(dB re 1µPa²/Hz)", na.value = "grey95") +
    theme_custom +
    theme(legend.position = "bottom", 
          legend.title = element_text())+
    labs(title = "<b>Long-Term Spectral Average (LTSA).</b> <i style='font-size:14pt'> Acoustic energy distribution across deployment,12-hour resolution and TOB frequency bins <br/> clipped to 10-9500 Hz.</i>")+
    labs(x = "Date", y = "Frequency (Hz)")
} else {
  cat("LTSA file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/LTSA.jpg", last_plot(), height = 10, width = 15, units = "cm")
```

\vspace{0.1in}

```{r loud_event, fig.width=12, fig.height=6, out.width="100%"}
#spl_file <- file.path(output_path, "all_spl.csv")
loud_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/loud_events.csv"))

if (file.exists(loud_file)) {
  loud <- read_csv(loud_file, show_col_types = FALSE)
  loud$DateTime <- as.POSIXct(loud$start_time)
  loud <- filter(loud,
                 #spectral_centroid_hz >=100,
                 delta_peak_dB >=10)
  
  loud_12h <- mutate(loud, DateTime12 = floor_date(DateTime, "24 hours"))%>%
    group_by(DateTime12)%>%
    summarise(ncount = n())

  count <- ggplot(data = loud_12h, aes(x = DateTime12, y = ncount))+
    geom_col(fill = "lightgrey", colour = "black", linewidth = 0.2, )+
    scale_x_datetime("Date", limits = c(start, end), expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
    scale_y_continuous("Events/day")+
    theme_custom+
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank())+
    labs(title = "<b>Broadband loud event detections.</b> <i style='font-size:14pt'> (Acoustic energy > 10dB above ambient levels) </i>")
  
  events <- ggplot(loud, aes(x = DateTime, y=spectral_centroid_hz, colour = delta_peak_dB, alpha = delta_peak_dB))+
    geom_point(size = 2, alpha = 0.8) +
    scale_y_log10("Spectral Centroid Frequency (Hz)")+
    scale_x_datetime("Date", limits = c(start, end), expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
    scale_colour_viridis_c("Delta SPL (dB)", limits = c(10,20))+
    theme_custom+
    theme(legend.position = "bottom",
          legend.title = element_text(vjust = 0.7))+
    labs(x = "Date", y = "SPL (dB re 1µPa)")
  
  grid.arrange(count, events,  ncol = 1, heights = c(0.4,1))
  
} else {
  cat("SPL file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/TOB_SPL.jpg", last_plot(), height = 10, width = 15, units = "cm")
```

\vspace{0.1in}

```{r spl_plot, fig.width=12, fig.height=7, out.width="100%"}
#spl_file <- file.path(output_path, "all_spl.csv")
spl_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/spl_timeseries_12h.csv"))

if (file.exists(spl_file)) {
  spl <- read_csv(spl_file, show_col_types = FALSE)
  spl$DateTime <- as.POSIXct(spl$DateTime)

  spl$band_hz <- ordered(spl$band_hz, levels = c(500, 200, 125, 63), labels = c("500Hz","200Hz","125Hz","63Hz"))

  ggplot(spl, aes(x = DateTime, y=mean, group = as.factor(band_hz), colour = as.factor(band_hz)))+
    geom_line(alpha = 0.8, linewidth = 0.5) +
    geom_point(size = 0.8, alpha = 0.8) +
    geom_ribbon(aes(ymin = mean-std, ymax = mean+std, fill = as.factor(band_hz)), alpha = 0.2, linewidth = 0.2)+
    facet_wrap(~as.factor(band_hz), ncol = 1)+
    scale_y_continuous("SPL (dB re 1 uPa)")+
    scale_x_datetime("Date", expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
    theme_custom+
    theme(legend.position = "none",
          strip.background =element_rect(fill="lightgrey"),
          strip.text = element_text(family = "Helvetica",size = 10))+
    labs(title = "<b>Sound Pressure Level (SPL) time series.</b> <i style='font-size:14pt'> (12-hourly averaged sound pressure levels in selected TOBs.) </i>")+
    labs(x = "Time", y = "SPL (dB re 1µPa)")
} else {
  cat("SPL file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/TOB_SPL.jpg", last_plot(), height = 10, width = 15, units = "cm")
```


\pagebreak

## Cetacean and Sonar Detections

\vspace{0.2in}

```{r cetacean_presence, fig.width=12, fig.height=8, out.width="100%"}
# Create timeline of detections and species pie chart

dol <- read.csv(list.files(paste0(data_path,"Processed_Sensors/FPod/S21/S21_R1"), pattern = "Dol TrainDuration.csv", full.names = T))
dol$Class <- "Dolphin"

nb <- read.csv(list.files(paste0(data_path,"Processed_Sensors/FPod/S21/S21_R1"), pattern = "NB TrainDuration.csv", full.names = T))
nb$Class <- "Porpoise"

sonar <- read.csv(list.files(paste0(data_path,"Processed_Sensors/FPod/S21/S21_R1"), pattern = "Sonar TrainDuration.csv", full.names = T))
sonar$Class <- "Sonar"

cet_all <- rbind(dol, nb, sonar)

cet_all$ChunkEnd <- as.POSIXct(cet_all$ChunkEnd, format = "%Y/%m/%d %H:%M", tz = "GMT")
cet_all <- filter(cet_all,
                  ChunkEnd>= start, ChunkEnd<=end)%>%
  distinct()
cet_all$Hour <- hour(cet_all$ChunkEnd)
cet_all$Day <- date(cet_all$ChunkEnd)
cet_all$Week <- week(cet_all$ChunkEnd)

cet_all_w <- group_by(cet_all, Day, Class)%>%
  summarise(TrainDuration = sum(Millisecs)/1000/60)

cet_timeline <- ggplot(cet_all_w, aes(x = Day, y = TrainDuration, fill = Class)) +
  geom_col(fill = "white") +
  geom_col(alpha = 0.9) +
  scale_fill_discrete(palette = c("#2f3d61","#462f61","#122b1f"))+
  scale_y_continuous(expand = c(0.03,0.0))+
  scale_x_datetime(expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
  theme_custom +
  theme(legend.position = "none",
        strip.background =element_rect(fill="lightgrey"),
        strip.text = element_text(family = "Helvetica",size = 10))+
  labs(title = "<b>Detections from FPod sensor.</b> <i style='font-size:14pt'> (Daily sum of train duration; Detection class: Hi) </i>")+
  labs(x = "Date", y = "Total train duration (min)")+
  facet_wrap(~Class, ncol = 1, scales = "free_y")

cet_timeline
```

\vspace{0.1in}

```{r cetacean_dielpatterns, fig.width=12, fig.height=5, out.width="100%"}
ndol <- sum(dol$DP10M)
cet_dial <- group_by(cet_all, Class, Hour)%>%
  summarise(mean = mean(Millisecs),
            sd = sd(Millisecs),
            sum = sum(Millisecs))%>%
  ungroup()%>%
  group_by(Class)%>%
  mutate(perc = 100*sum/sum(sum))

cet_rose <- ggplot(cet_dial, aes(x = Hour+0.5, y = perc, fill = Class))+
  geom_col(fill = "white") +
  geom_col(alpha = 0.9) +
  scale_fill_discrete(palette = c("#2f3d61","#462f61","#122b1f"))+
  scale_x_continuous(limits = c(0,24), breaks = seq(0,24, by = 1))+
  #scale_y_continuous(limits = c(0,25))+
  #geom_point(size = 3, colour = "#2f3d61") +
  coord_polar()+
  theme_custom +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        strip.background =element_rect(fill="lightgrey"),
        strip.text = element_text(family = "Helvetica",size = 10),
        aspect.ratio = 1)+
  labs(title = "<b>Diel pattern of detections from FPod sensor.</b> <i style='font-size:14pt'> (% of total train duration by hour. Detection class: Hi. Note varying y-axis) </i>")+
  labs(x = "Hour", y = "% Total train duration")+
  facet_wrap(~Class, nrow = 1, scales = "free_y")

cet_rose

```

\pagebreak

## Acoustic Tag Detections

```{r VR2AR_detections, fig.width=12, fig.height=5, out.width="100%"}
# Create timeline of detections

det <- read.csv('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/Processed_Sensors/VR2AR/S21/S21_R1/VR2AR-69 556544 2026-02-13 155019 DETECTS_noHeader.csv')
det$DateTime<- as.POSIXct(det$Device.Time..UTC.)

det <- filter(det,
              DateTime>=start, DateTime<=end)

det_timeline <- ggplot(det, aes(x = DateTime, y = as.factor(Serial.Number))) +
  geom_point(fill = "#2f3d61", alpha = 0.9) +
  scale_x_datetime(expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
  theme_custom +
  labs(title = "<b>Timeline of detections from VR2AR.</b> <i style='font-size:14pt'> </i>")+
  labs(x = "Date", y = "Serial")

det_timeline

```

***
