---
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=1.5cm"
output:
  pdf_document:
    toc: false
---

<style>
body { font-family: Helvetica, sans-serif; margin: 20px; }
.blue-panel { background-color: #e8f4f8; border: 2px solid #4a90e2; padding: 20px; margin: 20px 0; border-radius: 5px; }
h1, h2, h3 { color: #2c3e50; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%",
  dev = "png",        # force PNG instead of SVG (SVG is huge for rasters)
  dpi = 300
)

library(ggplot2)
library(readr)
library(tidyr)
library(dplyr)
library(scales)
library(gridExtra)
library(grid)
library(lubridate)
library(sf)
library(patchwork)
# Additional libraries for weather data
library(httr)
library(jsonlite)

# Set data path
data_path <- '/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/'

## Get station metadata
meta <- readxl::read_xlsx(paste0(data_path,"MetaData.xlsx"))

## Set station of interest
station <- 21

# Load spatial data
# Land polygon (coastline)
land_sf <- read_sf('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/GIS/Land/gadm36_0.shp')

# Stations point data (8 stations in a grid)
stations_sf <- read_sf("/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/GIS/s3_locations_wgs84.shp")
names(stations_sf)[1]<-"Station"

# Get coordinates of the focal station
focal_station <- stations_sf %>% filter(Station == station)

# Define bounding box (if not provided, use extent of all data)
bbox <- c(xmin = -8, xmax = -4, ymin = 49.8, ymax = 52.2)

```

```{r graphics, include=FALSE}

theme_custom <- theme_classic()+
  theme(
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.text = element_text(family = "Helvetica", size = 10),
    axis.title = element_text(family = "Helvetica", size = 12),
    plot.title = element_text(family = "Helvetica", size = 14, face=("bold")),
    panel.background = element_rect(fill = "white", colour = "NA"),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.grid = element_line(colour = "grey90", linewidth = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.direction = "horizontal"
  )

```


# Station S `r sprintf("%02d",station)` - Service Report


```{r title_page_layout, fig.width=12, fig.height=6, out.width="100%"}

# Create a simple title page with station info on left and placeholder for map on right
start <- meta$DateTime_deploy_UTC[meta$Station==paste0("S",station)]
end <- meta$DateTime_retrieve_UTC[meta$Station==paste0("S",station)]

# Station details
station_info <- data.frame(
  label = c("Station N:",
            "Deployment Period:",
            "Position:",
            "Depth:",
            "VR2AR Serial:",
            "FPOD Serial:",
            "Hydrophone Serial:",
            "Hydrophone duty cycle:"),
  value = c(paste0("S", sprintf("%02d",station)),
            paste0(date(start)," to ", date(end)),
            paste0(meta$Latitude_deploy[meta$Station==paste0("S", sprintf("%02d",station))],", ",meta$Longitude_deploy[meta$Station==paste0("S", sprintf("%02d",station))]),
            "120 m",
            meta$ARC_VR2AR[meta$Station==paste0("S", sprintf("%02d",station))],
            meta$FPOD[meta$Station==paste0("S", sprintf("%02d",station))],
            meta$Hydrophone[meta$Station==paste0("S", sprintf("%02d",station))],
            "50% on (15 min)"
  )
)

# Create text layout (left side)
left_plot <- ggplot() +
  geom_rect(aes(xmin = 0, xmax = 1, ymin = 0.82, ymax = 1.0),
            fill = "#4472C4", color = "black", linewidth = 0.5) +
  geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = 0.82),
            fill = "white", color = "black", linewidth = 0.5) +
  geom_text(aes(x = 0.1, y = 0.9, label = "Station Details"),
            size = 6, fontface = "bold", hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.75,
                label = paste(station_info$label[1], station_info$value[1], sep = " ")),
            size = 4, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.65,
                label = paste(station_info$label[2], station_info$value[2], sep = " ")),
            size = 4, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.55,
                label = paste(station_info$label[3], station_info$value[3], sep = " ")),
            size = 4, hjust = 0) +
  geom_text(aes(x = 0.1, y = 0.45,
                label = paste(station_info$label[4], station_info$value[4], sep = " ")),
            size = 4, hjust = 0) +
    geom_text(aes(x = 0.1, y = 0.35,
                label = paste(station_info$label[5], station_info$value[5], sep = " ")),
            size = 4, hjust = 0) +
    geom_text(aes(x = 0.1, y = 0.25,
                label = paste(station_info$label[6], station_info$value[6], sep = " ")),
            size = 4, hjust = 0) +
    geom_text(aes(x = 0.1, y = 0.15,
                label = paste(station_info$label[7], station_info$value[7], sep = " ")),
            size = 4, hjust = 0) +
      geom_text(aes(x = 0.1, y = 0.05,
                label = paste(station_info$label[8], station_info$value[8], sep = " ")),
            size = 4, hjust = 0) +
  xlim(0, 1) + ylim(0, 1) +
  theme_void() +
  theme(plot.background = element_rect(fill = "white", colour = NA))

# Create placeholder map (right side) - simple scatter plot as placeholder
map_plot <- ggplot() +
  # Land
  geom_sf(data = land_sf, fill = "gray80", color = "gray50", size = 0.3) +
  # All stations (small points)
  geom_sf(data = stations_sf, shape = 21, fill = "white", color = "black", size = 2) +
  # Focal station (larger, different symbol)
  geom_sf(data = focal_station, shape = 21, fill = "red", color = "black", size = 4) +
  # Station label
  geom_sf_text(data = focal_station, aes(label = Station),
               nudge_x = 0.07, nudge_y = 0.07, size = 5, fontface = "bold") +
  # Crop to bounding box
  coord_sf(xlim = c(bbox["xmin"], bbox["xmax"]),
           ylim = c(bbox["ymin"], bbox["ymax"]),
           expand = FALSE) +
  theme_custom+
  theme(
    #plot.background = element_rect(fill = "#e8f4f8", colour = "grey70", linewidth = 1),
    panel.background = element_rect(fill = "#e8f4f8", color = "black", linewidth = 1.5),
    panel.grid = element_line(colour = "grey90"),
    plot.margin = margin(0, 0, 0, 0)
  )+
  labs(title = "Station location", x = "Longitude", y = "Latitude")

# Combine plots side by side
grid.arrange(left_plot, map_plot, ncol = 2, widths = c(0.4, 0.6))
```

\vspace{0.1in}

## Environmental Conditions

```{r ctd_current_wind, fig.width=14, fig.height=8, out.width="100%"}

# Download weather data
base_url <- "https://archive-api.open-meteo.com/v1/archive"

# Build URL
url <- paste0(base_url,
              "?latitude=", focal_station$Y,
              "&longitude=", focal_station$X,
              "&start_date=", as.character(date(start)),
              "&end_date=", as.character(date(end)),
              "&hourly=temperature_2m,precipitation,wind_speed_10m,wind_direction_10m",
              "&timezone=GMT")

# Make request
response <- GET(url)

if (status_code(response) == 200) {
  data <- fromJSON(content(response, "text"))

  # Create data frame
  weather_df <- data.frame(
    DateTime = as.POSIXct(data$hourly$time, format = "%Y-%m-%dT%H:%M",tz = "GMT"),
    temperature = data$hourly$temperature_2m,
    rainfall = data$hourly$precipitation,
    wind_speed = data$hourly$wind_speed_10m,
    wind_direction = data$hourly$wind_direction_10m
  )
}
# Summarise at 12 hour intervals
weather_df <- weather_df %>%
  mutate(DateTime = floor_date(DateTime, "12 hours")) %>%
  group_by(DateTime) %>%
  summarise(
    temperature_mean    = mean(temperature),
    temperature_sd      = sd(temperature),
    wind_speed_mean     = mean(wind_speed),
    wind_speed_sd       = sd(wind_speed),
    wind_direction_mean = mean(wind_direction),
    wind_direction_sd   = sd(wind_direction)
  )


# Plot Wind Speed data
p1 <- ggplot(data = weather_df, aes( x = DateTime, y = wind_speed_mean))+
  geom_ribbon(aes(ymin = 0, ymax = wind_speed_mean), fill = "#001744", colour = "NA", alpha = 0.7)+
  geom_line(colour = "#001744", linewidth = 0.5)+
  scale_y_continuous("Wind speed (m/s)", expand = c(0,0))+
  scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "Wind Speed")

# Plot wind direction data
p2 <- ggplot(data = weather_df, aes( x = DateTime, y = wind_direction_mean))+
  geom_line(colour = "#001744", linewidth = 0.5)+
  scale_y_continuous("Wind direction", expand = c(0,0))+
  scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "Wind Direction")

# Get current data from POLPRED
current_data <- read.csv('/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/POLPRED/S21_14oct25_16Jan26.csv', skip = 4)
current_data$time <- as.POSIXct(current_data$Date.Time..UTC., format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC")

current_data <- filter(current_data,
                       time>= start, time <= end)
names(current_data)[3] <- "current_speed"

cmean <- current_data %>%
  mutate(DateTime = floor_date(time, "12 hours")) %>%
  group_by(DateTime) %>%
  summarise(
    current_speed_mean    = mean(current_speed),
    current_speed_sd    = sd(current_speed)
  )

# Plot current data
p3 <- ggplot(current_data, aes(x = time, y = current_speed)) +
  geom_point(size = 1, colour = "#1b9e77", alpha = 0.5) +
  geom_ribbon(data= cmean, aes(x = DateTime, ymin = current_speed_mean-current_speed_sd, ymax = current_speed_mean+current_speed_sd), fill = "grey80", alpha = 0.5, inherit.aes = F)+
  geom_line(data= cmean, aes(x = DateTime, y = current_speed_mean), colour = "#1b9e77", linewidth = 1.5, inherit.aes = F)+
   scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "Bottom Current Speed", x = "Time (hours)", y = "Speed (m/s)")

# Get temeprature data from VR2AR
# Read skipping the first 4 header rows, no column names
raw <- read_csv("/Users/f.garzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Data/Processed_Sensors/VR2AR/S21/S21_R1/VR2AR-69 556544 2026-02-13 155019 SENSORS.csv",
                skip = 4,
                col_names = FALSE,
                show_col_types = FALSE)

# Keep only TEMP rows and select relevant columns
temp_df <- raw %>%
  filter(X1 == "TEMP") %>%
  select(
    record_type  = X1,
    datetime     = X2,
    model        = X6,
    serial       = X7,
    temp_ambient = X8
  ) %>%
  mutate(
    datetime     = dmy_hm(datetime),
    temp_ambient = as.numeric(temp_ambient),
    DateTime = floor_date(datetime, "12 hours")) %>%
  filter(DateTime >= start,
         DateTime <= end)%>%
  group_by(DateTime) %>%
  summarise(
    temperature_mean    = mean(temp_ambient),
    temperature_sd      = sd(temp_ambient)
  )

p4 <- ggplot(temp_df, aes(x = DateTime)) +
  geom_ribbon(aes(ymin = temperature_mean-temperature_sd, ymax = temperature_mean+temperature_sd), fill = "grey80", alpha = 0.5)+
  geom_line(aes(y = temperature_mean), colour = "#d17432", linewidth = 1.5)+
  #geom_ribbon(data = weather_df, aes(ymin = temperature_mean-temperature_sd, ymax = temperature_mean+temperature_sd), fill = "grey80", alpha = 0.5)+
  #geom_line(data = weather_df, aes(y = temperature_mean))+
   scale_x_datetime("Date", breaks = "10 days", expand = c(0,0))+
  theme_custom+
  labs(title = "Bottom water temperature", x = "Date", y = "°C")

grid.arrange(p1, p2, p3, p4, ncol = 1)
```

Wind data: OpenMeteo. Current data : POLPRED. Temperature data: VR2AR sensor.

\pagebreak

## Soundscape and Noise

\vspace{0.1in}

```{r quantiles_plot, fig.width=11, fig.height=4.5, out.width="100%"}
#quant_file <- file.path(output_path, "all_quantiles.csv")
quant_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/quantile_spectral_density_12h.csv"))

if (file.exists(quant_file)) {
  q <- read_csv(quant_file, show_col_types = FALSE)

  ggplot(q, aes(x = TOB, y = SPL, color = as.factor(percentile))) +
    geom_line(linewidth = 0.9) +
    scale_x_log10(
      breaks = c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000, 20000, 50000),
      labels = c("10","20","50","100","200","500","1k","2k","5k","10k","20k","50k"),
      expand = c(0,0)
    ) +
    scale_color_manual(
      values = c("1" = "#fee5d9", "5" = "#fcae91", "25" = "#fb6a4a",
                 "75" = "#de2d26", "95" = "#a50f15","99" = "darkred", "50" = "#006837"),
      labels = c("1" = "1st %ile", "5" = "5th %ile", "25" = "25th %ile", "50" = "Median",
                 "75" = "75th %ile", "95" = "95th %ile", "99" = "99th %ile")
    ) +
    theme_custom +
    theme(legend.position = "right", legend.direction = "vertical")+
    labs(title = "Spectral Quantiles", x = "Frequency (Hz)", y = "Power (dB re 1µPa²/Hz)")
} else {
  cat("Quantiles file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/Quatiles.jpg", last_plot(), height = 10, width = 15, units = "cm")
```
Distribution of acoustic energy across TOB frequencies.

\vspace{0.1in}

```{r ltsa_plot, fig.width=11, fig.height=5, out.width="100%"}
#ltsa_file <- file.path(output_path, "all_ltsa.csv")
ltsa_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/ltsa_12h.csv"))

if (file.exists(ltsa_file)) {
  ltsa <- read_csv(ltsa_file, guess_max = 10000, show_col_types = FALSE)

  ltsa$DateTime <- as.POSIXct(ltsa$DateTime, tz = "GMT")

  ggplot(ltsa, aes(x = DateTime, y = TOB, fill = SPL)) +
    geom_raster(interpolate =T) +
    geom_vline(xintercept=as.POSIXct(c("2025-11-01 00:00:00",
                                       "2025-12-01 00:00:00",
                                       "2026-01-01 00:00:00"), tz = "UTC"),
               colour = "white", linetype = 2)+
    scale_y_log10(limits = c(10,9400), expand = c(0,0),
                  breaks = c(10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 9000),
                  labels = c("10","20","50","100","200","500","1k","2k","5k","9k")) +
    scale_x_datetime(expand = c(0,0), breaks = "15 days", date_labels = "%b %d")+
    #scale_y_log10()+
    scale_fill_viridis_c(name = "Power\n(dB re 1µPa²/Hz)", na.value = "grey95") +
    theme_custom +
    theme(legend.position = "right", legend.direction = "vertical")+
    labs(title = "Long-Term Spectral Average (LTSA)", x = "Time", y = "Frequency (Hz)")
} else {
  cat("LTSA file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/LTSA.jpg", last_plot(), height = 10, width = 15, units = "cm")
```
Acoustic energy distribution across deployment, with 12-hour temporal resolution and TOB frequency bins clipped to 10-9500 Hz.

\vspace{0.1in}

```{r spl_plot, fig.width=11, fig.height=10, out.width="100%"}
#spl_file <- file.path(output_path, "all_spl.csv")
spl_file <- file.path(paste0(data_path,"Processed_Sensors/SoundTrap/S21/S21_R1/spl_timeseries_12h.csv"))

if (file.exists(spl_file)) {
  spl <- read_csv(spl_file, show_col_types = FALSE)
  spl$DateTime <- as.POSIXct(spl$DateTime)

  spl$band_hz <- ordered(spl$band_hz, levels = c(500, 200, 125, 63), labels = c("500Hz","200Hz","125Hz","63Hz"))

  ggplot(spl, aes(x = DateTime, y=mean, group = as.factor(band_hz), colour = as.factor(band_hz)))+
    geom_line(alpha = 0.8, linewidth = 0.5) +
    geom_point(size = 0.8, alpha = 0.8) +
    geom_ribbon(aes(ymin = mean-std, ymax = mean+std, fill = as.factor(band_hz)), alpha = 0.2, linewidth = 0.2)+
    facet_wrap(~as.factor(band_hz), ncol = 1)+
    scale_y_continuous("SPL (dB re 1 uPa)")+
    scale_x_datetime("Date", expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
    theme_custom+
    theme(legend.position = "none")+
    labs(title = "Sound Pressure Level (SPL) Time Series", x = "Time", y = "SPL (dB re 1µPa)")
} else {
  cat("SPL file not found")
}

#ggsave("/Users/francescogarzon/Library/CloudStorage/OneDrive-UniversityofExeter/Projects/S3 Project - Documents/Figures/S4Station/TOB_SPL.jpg", last_plot(), height = 10, width = 15, units = "cm")
```
12-hourly averaged sound pressure levels in key third-octave bands.

\pagebreak

## Cetacean Detections

Cetacean presence is expressed in "Detection-Positive 10-Minute intervals" (DP10M). These intervals represent the number of 10-minute intervals per week in which vocalisations were detected.

\vspace{0.2in}

```{r cetacean_presence, fig.width=14, fig.height=8, out.width="100%"}
# Create timeline of detections and species pie chart

dol <- read.csv(list.files(paste0(data_path,"Processed_Sensors/FPod/S21/S21_R1"), pattern = "Dol DP10M.csv", full.names = T))
dol$ChunkEnd <- as.POSIXct(dol$ChunkEnd, format = "%d/%m/%Y %H:%M", tz = "GMT")
dol <- filter (dol,
               ChunkEnd>= start, ChunkEnd<=end)
dol$Hour <- hour(dol$ChunkEnd)
dol$Day <- date(dol$ChunkEnd)
dol$Week <- week(dol$ChunkEnd)

#dol <- filter(dol, Day >= as.POSIXct("2025-10-14 00:00:00"), Day <= as.POSIXct("2025-10-17 00:00:00"))

nb <- read.csv(list.files(paste0(data_path,"Processed_Sensors/FPod/S21/S21_R1"), pattern = "NB DP10M.csv", full.names = T))
nb$ChunkEnd <- as.POSIXct(nb$ChunkEnd, format = "%d/%m/%Y %H:%M", tz = "GMT")
nb <- filter(nb,
             ChunkEnd>= start, ChunkEnd<=end)
nb$Hour <- hour(nb$ChunkEnd)
nb$Day <- date(nb$ChunkEnd)
nb$Week <- week(nb$ChunkEnd)
#nb <- filter(nb, Day >= as.POSIXct("2025-10-14 00:00:00"), Day <= as.POSIXct("2025-10-17 00:00:00"))

dol_w <- group_by(dol, Day)%>%
  summarise(DP10M_sum = sum(DP10M))
nb_w <- group_by(nb, Day)%>%
  summarise(DP10M_sum = sum(DP10M))

dol_timeline <- ggplot(dol_w, aes(x = Day, y = DP10M_sum)) +
  geom_col(fill = "white") +
  geom_col(fill = "#2f3d61", alpha = 0.9) +
  scale_y_continuous(expand = c(0.03,0.0))+
  scale_x_datetime(expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
  theme_custom +
  labs(title = "Dolphin Detection-positive 10 minute intervals (DP10M)", x = "Date", y = "# DP10M")

nb_timeline <- ggplot(nb_w, aes(x = Day, y = DP10M_sum)) +
  geom_col(fill = "white") +
  geom_col(fill = "#462f61", alpha = 0.9) +
  scale_y_continuous(expand = c(0.03,0.0))+
  scale_x_datetime(expand = c(0,0), breaks = "10 days", date_labels = "%b %d")+
  theme_custom +
  labs(title = "Porpoise Detection-positive 10 minute intervals (DP10M)", x = "Date", y = "# DP10M")


grid.arrange(dol_timeline, nb_timeline, ncol = 1, heights = c(0.5, 0.5))
```

\vspace{0.2in}

```{r cetacean_dielpatterns, fig.width=8, fig.height=5, out.width="100%"}
ndol <- sum(dol$DP10M)
dol_dial <- group_by(dol, Hour)%>%
  summarise(mean = mean(DP10M),
            sd = sd(DP10M),
            sum = sum(DP10M),
            perc = 100 * sum/ndol)

dol_dial2h <- mutate(dol_dial,
                     Hour2 = as.integer(as.character(cut(Hour,
                                                         breaks = seq(0,24, by  =2),
                                                         right = F,
                                                         labels = seq(0,23, by  =2))))
                     )%>%
  group_by(Hour2)%>%
  summarise(perc = sum(perc))

nnb <- sum(nb$DP10M)
nb_dial <- group_by(nb, Hour)%>%
  summarise(mean = mean(DP10M),
            sd = sd(DP10M),
            sum = sum(DP10M),
            perc = 100 * sum/nnb)

nb_dial2h <- mutate(nb_dial,
                     Hour2 = as.integer(as.character(cut(Hour,
                                                         breaks = seq(0,24, by  =2),
                                                         right = F,
                                                         labels = seq(0,23, by  =2))))
                     )%>%
  group_by(Hour2)%>%
  summarise(perc = sum(perc))


dol_rose <- ggplot(dol_dial2h, aes(x = Hour2, y = perc))+
  geom_col(fill = "#2f3d61", alpha = 0.9, just = 0) +
  scale_x_continuous(limits = c(0,24), breaks = seq(0,24, by = 2))+
  scale_y_continuous(limits = c(0,25))+
  #geom_point(size = 3, colour = "#2f3d61") +
  coord_polar()+
  theme_custom +
  labs(title = "% Dolphin DP10M", x = "Date", y = "% DP10M")

nb_rose <- ggplot(nb_dial2h, aes(x = Hour2, y = perc))+
  geom_col(fill = "#462f61", alpha = 0.9, just = 0) +
  scale_x_continuous(limits = c(0,24), breaks = seq(0,24, by = 2))+
  scale_y_continuous(limits = c(0,25))+
  #geom_point(size = 3, colour = "#2f3d61") +
  coord_polar()+
  theme_custom +
  labs(title = "% Porpoise DP10M", x = "Hour", y = "% DP10M")

grid.arrange(dol_rose, nb_rose, ncol = 2)

```
\vspace{0.3in}

***

**Report Generated:** `r format(Sys.Date(), "%B %d, %Y")`
